Simple models of deliberated preference

### 2018-01-28 simple.tex
Intro with phi’s

TODO do everything with only one relation. Define for each prop: $\prop^{s}$ and $\prop^p$, and define $¬(\prop^s) = (¬\prop)^p$, and we can do everything with $(\ar, \ar') \ileadsto \prop^p$ (defined as $\ileadstopst \prop$) and $(\ar, \ar') \ileadsto \prop^s$ (defined as $\ileadstosst \prop$).

Actually we will use only the $\ileadstopst$ and $\ileadstosst$ ternary relations from now on (and their negations $\ileadstose$ and $\ileadstope$), together with the property (A1) that $\ileadstopst \subseteq \ileadstope$
\footnote{Equivalently: $\ileadstosst \subseteq \ileadstose$, $(\ar, \ar') \ileadstose ¬\prop ∨ (\ar, \ar') \ileadstope \prop$, $\ileadstopst \prop ⇒ ¬ (\ileadstosst ¬\prop)$, $\ileadstosst ¬\prop ⇒ ¬ (\ileadstopst \prop)$.} 
and (A2) that $\ileadstosst \subseteq \ileadstopst$
\footnote{Equivalently: $\ileadstose \subseteq \ileadstope$.}, 
the pair $(\ar, \ar')$ being understood as unordered (thus not different than $(\ar', \ar)$).

We also define $\Phi = \alltopic × \set{\text{possible}, \text{sure}}$, and $\forall \phi = (\prop, \text{possible}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadstopst \prop$ and $\forall \phi = (\prop, \text{sure}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadstosst \prop$. We define the negation operator $¬$ over $\Phi$ as: $¬(\prop, \text{possible}) = (¬\prop, \text{sure})$ and $¬(\prop, \text{sure}) = (¬\prop, \text{possible})$. It follows that $¬¬\phi = \phi$. We write $(\ar, \ar') \nileadstost \phi$ for $¬[(\ar, \ar') \ileadstost \phi]$.
\footnote{From (A1) and (A2) we can write, using this notation, $(\ar, \ar') \ileadstost \phi ⇒ (\ar, \ar') \nileadstost ¬\phi$.}

$\ar$ is $\phi\text{-dec} ⇔ \forall \ar': (\ar', \ar) \ileadstost \prop$. We also say that $\ar$ is $\propsuredec$ or $\proppossdec$ when $\phi=(\prop, \text{sure})$ or $\phi=(\prop, \text{possible}$.
\footnote{$\propsuredec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstosst \prop$, $\proppossdec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstopst \prop$.}

We can now define $\iposs, \isure \subseteq \alltopic$:
\begin{itemize}
	\item $\prop \in \iposs ⇔ \exists \ar \ \proppossdec$. 
\footnote{\label{ft:notposs}$\prop \notin \iposs ⇔ \forall \ar, \exists \ar' \suchthat (\ar', \ar) \ileadstose ¬\prop$.}
	\item $\prop \in \isure ⇔ \exists \ar \ \propsuredec$. 
\footnote{\label{ft:notsure}$¬\prop \notin \isure ⇔ \forall \ar, \exists \ar' \suchthat (\ar', \ar) \ileadstope \prop$.}
\end{itemize}

@@

\begin{theorem}
	$\prop \in \isure ⇒ ¬\prop \notin \iposs$.
\end{theorem}
This is because some $\ar^+$ is $\propsuredec$, thus given any $\ar^−$, $(\ar^+, \ar^−) \ileadstosst \prop$ hence $(\ar^+, \ar^−) \ileadstose \prop$ (using A1).

\begin{theorem}
	$\prop \in \isure ⇒ \prop \in \iposs$.
\end{theorem}
The proof follows from (A2).

\begin{definition}[Clear-cut]
	A situation is $\prop$-clear-cut iff $\prop \notin \iposs ⇒ ¬\prop \in \isure$. A situation is $(\prop, ¬\prop)$-clear-cut iff $\prop \notin \isure ∧ ¬\prop \notin \isure ⇒ \prop \in \iposs ∧ ¬\prop \in \iposs$.
\footnote{$\prop\text{-CC} ∧ ¬\prop\text{-CC} ⇔ (¬\prop \in \isure ∨ \prop \in \iposs) ∧ (\prop \in \isure ∨ ¬\prop \in \iposs) ⇔ (¬\prop \in \isure ∧ \prop \in \isure) ∨ (¬\prop \in \isure) ∨ (\prop \in \isure) ∨ (\prop \in \iposs ∧ ¬\prop \in \iposs) ⇔ (\prop, ¬\prop)\text{-CC}$.}
\footnote{$\overline{\prop\text{-CC}} ∧ \overline{¬\prop\text{-CC}} ⇔ (\prop, ¬\prop)$-unstability.}
\footnote{Example of $\overline{¬\prop\text{-CC}} ∧ \prop \in \iposs$ (thus with $\prop\text{-CC}$ and no $(\prop, ¬\prop)$-unstability): $\forall (\ar, \ar'), (\ar, \ar') \ileadstose \prop, (\ar, \ar') \ileadstope ¬\prop, (\ar, \ar') \ileadstopst \prop$. In such a situation, it is unclear whether $\prop$ holds rather than $¬\prop$ or both $\prop$ and $¬\prop$.}
\footnote{Hence (with possible meaning merely possible and not sure), for a given pair $(t, ¬t)$, either both are possible, or one is sure, or only $¬\prop$-CC and $¬t \in \iposs$ and $t \notin \iposs$, or only $\prop$-CC and $t \in \iposs$ and $¬t \notin \iposs$, or no CC and $(\prop, ¬\prop)$-unstable.}
\footnote{Something like Axiom JU should be sufficient for CC. We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}
\end{definition}

@@

Notation. Let $\mleadstoinv(\alltopic) \subseteq \clargs$ denote the subset of arguments supporting propositions. 
$\mbeats(\args)$: arguments that $\args$ attacks, $\ar_1 \in \mbeats(\args) ⇔ \exists \ar_2 \in \args \suchthat \ar_2 \mbeats \ar_1$; $\mbeatsinv(\ar)$: arguments attacking $\ar$, $\ar_2 \in \mbeatsinv(\ar) ⇔ \ar_2 \mbeats \ar$.

>>>>

Actually we will use only the $\ileadstopst$ and $\ileadstosst$ ternary relations from now on (and their negations $\ileadstose$ and $\ileadstope$), together with the property (A1) that $\ileadstopst \subseteq \ileadstope$
\footnote{Equivalently: $\ileadstosst \subseteq \ileadstose$, $(\ar, \ar') \ileadstose ¬\prop ∨ (\ar, \ar') \ileadstope \prop$, $\ileadstopst \prop ⇒ ¬ (\ileadstosst ¬\prop)$, $\ileadstosst ¬\prop ⇒ ¬ (\ileadstopst \prop)$.} 
and (A2) that $\ileadstosst \subseteq \ileadstopst$
\footnote{Equivalently: $\ileadstose \subseteq \ileadstope$.}, 
the pair $(\ar, \ar')$ being understood as unordered (thus $(\ar, \ar') \ileadstosst \prop ⇔ (\ar', \ar) \ileadstosst \prop$, and similarly for $\ileadstopst$).

We also define $\Phi = \alltopic × \set{\text{possible}, \text{sure}}$, and $\forall \phi = (\prop, \text{possible}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadstopst \prop$ and $\forall \phi = (\prop, \text{sure}) \in \Phi: (\ar, \ar') \ileadstost \phi ⇔ (\ar, \ar') \ileadstosst \prop$. We define the negation operator $¬$ over $\Phi$ as: $¬(\prop, \text{possible}) = (¬\prop, \text{sure})$ and $¬(\prop, \text{sure}) = (¬\prop, \text{possible})$. It follows that $¬¬\phi = \phi$. We write $(\ar, \ar') \nileadstost \phi$ for $¬[(\ar, \ar') \ileadstost \phi]$ and define $(\ar, \ar') \ileadstoe \phi$ as equivalent to $(\ar, \ar') \nileadstost ¬\phi$.
\footnote{\label{ft:sttoe}From (A1) and (A2) we can write, using this notation, $(\ar, \ar') \ileadstost \phi ⇒ (\ar, \ar') \ileadstoe \phi$.}

$\ar$ is $\phi\text{-dec} ⇔ \forall \ar': (\ar', \ar) \ileadstost \phi$. 
\footnote{Equivalently: $\ar \ \phidec ⇔ \exists \ar \suchthat [\ileadstostinv(\phi)](\ar) = \allargs$.}
\footnote{We also say that $\ar$ is $\propsuredec$ or $\proppossdec$ when $\phi=(\prop, \text{sure})$ or $\phi=(\prop, \text{possible})$.
$\propsuredec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstosst \prop$, $\proppossdec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstopst \prop$.}

We can now define $\iPhi \subseteq \Phi$: $\phi \in \iPhi ⇔  \exists \ar \ \phi\text{-dec}$.
\footnote{Equivalently: $\iPhi = \bigcup_{\ar \in \allargs} \bigcap{\ar' \in \allargs} \ileadstost(\ar', \ar)$.}
\footnote{$\phi \in \iPhi ⇔ \exists \ar \suchthat \allargs × \{\ar\} \ileadstost \phi$.}
\footnote{\label{ft:notposs}$\prop \in \iposs ⇔ \exists \ar \ \proppossdec$ and $\prop \notin \iposs ⇔ \forall \ar, \exists \ar' \suchthat (\ar', \ar) \ileadstose ¬\prop$.}
\footnote{\label{ft:notsure}$\prop \in \isure ⇔ \exists \ar \ \propsuredec$ and $¬\prop \notin \isure ⇔ \forall \ar, \exists \ar' \suchthat (\ar', \ar) \ileadstope \prop$.}

@@

\begin{theorem}
	$\phi \in \iPhi ⇒ ¬\phi \notin \iPhi$.
\end{theorem}
This is because some $\ar^+$ is $\phidec$, thus given any $\ar^−$, $(\ar^+, \ar^−) \ileadstost \phi$ hence $(\ar^+, \ar^−) \ileadstoe \phi$ (using \cref{ft:sttoe}).

\begin{theorem}
	$(\prop, \text{sure}) \in \iPhi ⇒ (\prop, \text{possible}) \in \iPhi$.
\end{theorem}
The proof follows from (A2).

\begin{definition}[Clear-cut]
	A situation is $\prop$-clear-cut iff $(\prop, \text{sure}) \notin \iPhi ⇒ (¬\prop, \text{possible}) \in \iPhi$. A situation is $(\prop, ¬\prop)$-clear-cut iff $\prop \notin \isure ∧ ¬\prop \notin \isure ⇒ \prop \in \iposs ∧ ¬\prop \in \iposs$, equivalently, $\phi \notin \iPhi ⇔ ¬\phi \in \iPhi$.
\footnote{$\prop\text{-CC} ∧ ¬\prop\text{-CC} ⇔ (¬\prop \in \isure ∨ \prop \in \iposs) ∧ (\prop \in \isure ∨ ¬\prop \in \iposs) ⇔ (¬\prop \in \isure ∧ \prop \in \isure) ∨ (¬\prop \in \isure) ∨ (\prop \in \isure) ∨ (\prop \in \iposs ∧ ¬\prop \in \iposs) ⇔ (\prop, ¬\prop)\text{-CC}$.}
\footnote{$\overline{\prop\text{-CC}} ∧ \overline{¬\prop\text{-CC}} ⇔ (\prop, ¬\prop)$-unstability.}
\footnote{Example of $\overline{¬\prop\text{-CC}} ∧ \prop \in \iposs$ (thus with $\prop\text{-CC}$ and no $(\prop, ¬\prop)$-unstability): $\forall (\ar, \ar'), (\ar, \ar') \ileadstose \prop, (\ar, \ar') \ileadstope ¬\prop, (\ar, \ar') \ileadstopst \prop$. In such a situation, it is unclear whether $\prop$ holds rather than $¬\prop$ or both $\prop$ and $¬\prop$.}
\footnote{Hence (with possible meaning merely possible and not sure), for a given pair $(t, ¬t)$, either both are possible, or one is sure, or only $¬\prop$-CC and $¬t \in \iposs$ and $t \notin \iposs$, or only $\prop$-CC and $t \in \iposs$ and $¬t \notin \iposs$, or no CC and $(\prop, ¬\prop)$-unstable.}
\footnote{Something like Axiom JU should be sufficient for CC. We say $(\ar, \ar')$ is $\prop$-unstable iff $(\ar, \ar') \ileadstose \prop ∧ (\ar, \ar') \ileadstose ¬\prop$. JU: If $(\ar, \ar')$ is $\prop$-unstable, then $\exists \ar_2 \in \allargs$ such that for $(\ar_1, \ar_0) = (\ar, \ar')$ or for $(\ar_1, \ar_0) = (\ar', \ar)$, $\ar_2$ attacks decisively $\ar_1$, where we mean by this that $\ar_2$ combined with $\ar_0$ is decisive. Means that if not CC, then not JU. Hence, infinite discussion happens, or irreducible unstability. Or, the memory buffer of $i$ is too short to hold adequate reinstated arguments, in which case our approach based on decisive arguments is not adequate.}
\end{definition}

@@

Notation. Let $\mleadstoinv(\alltopic) \subseteq \clargs$ denote the subset of arguments supporting propositions. 
$\mbeats(\args)$: arguments that $\args$ attacks, $\ar_1 \in \mbeats(\args) ⇔ \exists \ar_2 \in \args \suchthat \ar_2 \mbeats \ar_1$; $\mbeatsinv(\ar)$: arguments attacking $\ar$, $\ar_2 \in \mbeatsinv(\ar) ⇔ \ar_2 \mbeats \ar$.  We write $\ar \mleadsto (\prop, \text{possible})$ for $\ar \mleadsto \prop$ and $\ar \mleadsto (\prop, \text{sure})$ for $\ar \mleadsto \prop ∧ (¬\prop \notin \mleadsto(\allargs))$.
